The Latent Constructive Rewrite Model \\ or \\ The Latent Conceptual Grammar Model

### 2014-10-09 drafts/cfp/draft.tex
Changes to aaai symposium draft.

\subsection{Related Work}

\section{The Model}
In this section we propose a hierarchical probabilistic model for
learning distributions over string relations. Suppose we are given
some positive instances $y_1, \dots, y_M$ of relations $R_1, \dots,
R_N$. Our model assumes that each positive instance
$y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent stochastic grammar
(LSG), $\mathcal{K}$. $\mathcal{K}$ consists of some fixed number $K$
of clauses (i.e. rules) $C_1, \dots, C_K$ over the $N$ observable
relations plus a fixed number of $H$ hidden relations $T_1, \dots,
T_H$. The \emph{architecture} of the model is the exact specification
of the parameters $K$ and $H$ and the clauses $mathcal{C}$. To each
clause $C_i$ is associated a positive real weights $\theta_i$.

>>>>

\subsection{Related Work}
This work borrows heavily

\section{The Model}
An LCG is a hierarchical Bayesian model of string tuples. It is an
extension of the Hierarhical Dirichlet PCFG model (HD-PCFG) to
Probabilistic Range Concatenation Grammars (PRCGs). Whereas the
HD-PCFG is a distribution over strings, an LCG is a distribution over
string tuples belonging to one of a set of observable
relations.

Range Concatenation Grammars (RCGs) are a class of string grammars
that represent all and only those languages that can be parsed in time
polynomial in the length of the target
string~\cite{boullier2005range}. An RCG $G=(N, T, V, P, S)$ is a
5-tuple where $N$ is a finite set of predicate symbols, $T$ is a set
of terminal symbols, $V$ is a set of variable symbols, P is a finite
set of $M \geq 0$ clauses of the form $\psi_0 \rightarrow \psi_1 \dots
\psi_M$, and $S \in N$ is the \emph{axiom}. Each $\psi_m$ is a term of
the form $A(\alpha_1, \dots, \alpha_{\mathcal{A}(A)}$, where $A \in
N$, $\mathcal{A}(A)$ is the arity of $A$, and each $\alpha_i \in (T
\cup V)^*$ is an argument of $\psi_m$. We call the left hand side term
of any clause the \emph{head} of that clause and its predicate symbol
is the \emph{head predicate}.

\textbf{Example:} The following is an RCG for the 3-copy language
$\{www | w \in {a, b}^*\}$:
\begin{alignat*}{3}
&S(XYZ) &&\rightarrow A(X, Y, Z)\\
&A(aX, aY, aZ) &&\rightarrow A(X, Y, Z)\\
&A(bX, bY, bZ) &&\rightarrow A(X, Y, Z)\\
&A(\epsilon, \epsilon, \epsilon) &&\rightarrow 
\end{alignat*}

A string $x$ is in the language defined by an RCG if one can
\emph{derive} $S(x)$. A derivation is a sequence of rewrite steps in
which substrings of the left hand side argument string are bound to
the variables of the head of some clause, thus determining the
arguments in the clause body. If a clause has no body terms, then its
head is derived; otherwise, its head is derived if its body clauses
are derived~\footnote{This description of the language of an RCG
  technically only holds for \emph{non-combinatory} RCGs, in which the
  arguments of body terms can only contain single variables. Since any
  \emph{combinatory} RCG can be converted into a non-combinatory RCG
  and we only consider non-combinatory RCGs here, this description
  suffices.}

We extend RCGs to Probabilistic Range Concatenation Grammars (PRCGs)
by annotating each clause $C_k \in P$ with probabilities $p_k$ such
that for all predicates $A \in N$ $\sum_{k:head(C_k)=A} p_k = 1$. A
PRCGs defines a distribution over strings $x$ by sampling from
derivations of $S(x)$ according to the product of probabilities of
clauses used in that derivation. This is a well defined distribution
as long as no probability mass is placed on derivations of infinite
length; in this paper, we only consider PRCGs with derivations of
finite length, so we need not worry about this requirement.

Now we can describe the model we introduce here, the Latent Concept
Grammar (LCG). We introduce the LCG as a machine learning model to
learn a distribution over instances of relations over strings. Unlike
a PRCG, which places a distribution over terms $(

RCGs consist of a set
of rewrite rules, called clauses, each of which takes a termwhich
operate on a set of terms (predicates applied to arguments) to produce
new terms. Each rewrite rule is of the form:

Suppose we are given some positive instances $y_1, \dots,
y_M$ of relations $R_1, \dots, R_N$. Our model assumes that each
positive instance $y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent
stochastic grammar (LSG), $\mathcal{K}$. $\mathcal{K}$ consists of
some fixed number $K$ of clauses (i.e. rules) $C_1, \dots, C_K$ over
the $N$ observable relations plus a fixed number of $H$ hidden
relations $T_1, \dots, T_H$. The \emph{architecture} of the model is
the exact specification of the parameters $K$ and $H$ and the clauses
$mathcal{C}$. To each clause $C_i$ is associated a positive real
weights $\theta_i$.
